# Load in necessary libraries

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(caret)
library(glue)
library(arrow)
library(stargazer)
library(treemap)

# set seed
set.seed(218)
```

# Read in all data

```{r}
# house data
static_house <- read_parquet("static_house_info.parquet")
# get each unique building ID
bldg_ids <- static_house %>% distinct(bldg_id) %>% .[[1]]
# take a random sample of building IDs
static_house %>% 
  group_by(in.county) %>% 
  sample_n(size = 5, replace = FALSE) %>% 
  ungroup() %>% 
  distinct(bldg_id) %>% 
  .[[1]] -> bldg_ids_sample
# get energy data for each house in the above sample
all_house_energy <- map(
  bldg_ids_sample, 
  function(x) read_parquet(
    glue(
      "https://intro-datascience.s3.us-east-2.amazonaws.com/SC-data/2023-houseData/{x}.parquet"
    )
  ) %>% 
    mutate(
      bldg_id = x,
      total_energy_consumption = rowSums(select(., ends_with("energy_consumption")))
    ),
  .progress = TRUE
) %>% 
  bind_rows()
# merge energy usage and static house datasets
all_house_energy %>% 
  left_join(static_house, by = "bldg_id") -> energy_usage
# get each unique county in house data
counties <- static_house %>% distinct(in.county) %>% .[[1]]
# get weather for each county
all_weather <- map(
  counties, 
  function(x) read_csv(
    glue(
      "https://intro-datascience.s3.us-east-2.amazonaws.com/SC-data/weather/2023-weather-data/{x}.csv"
    ),
    show_col_types = FALSE
  ) %>% 
    mutate(in.county = x),
  .progress = TRUE
) %>% 
  bind_rows()
# merge weather data with energy/static house data
energy_usage %>% 
  left_join(all_weather, by = c("in.county", "time" = "date_time")) -> all_data
# add month and time variables
all_data %>%
  rename("datetime" = "time") %>% 
  mutate(
    month = month(datetime, label = TRUE),
    time = format(datetime, "%H:%M:%S")   
  ) -> all_data
```

# Data visualizations

```{r}
# sqft
energy_usage %>% 
  group_by(bldg_id) %>% 
  summarise(mean_energy_usage = mean(total_energy_consumption),
            in.sqft = min(in.sqft)) %>% 
  ggplot(aes(x = as.factor(in.sqft), y = mean_energy_usage)) +
  geom_boxplot(color = "purple") +
  labs(
    title = "Square Feet vs. Total Energy Consumed",
    x = "Square Feet",
    y = "Total Energy Consumed (kWh)"
  ) +
  theme_minimal()

# cooling setpoint
energy_usage %>% 
  group_by(bldg_id) %>% 
  summarise(mean_energy_usage = mean(total_energy_consumption),
            in.cooling_setpoint = min(in.cooling_setpoint)) %>% 
  ggplot(aes(x = as.factor(in.cooling_setpoint), y = mean_energy_usage)) +
  geom_boxplot(color = "purple") +
  labs(
    title = "Cooling Setpoint vs. Total Energy Consumed",
    x = "Cooling Setpoint (F)",
    y = "Total Energy Consumed (kWh)"
  ) +
  theme_minimal()

# heating setpoint
energy_usage %>% 
  group_by(bldg_id) %>% 
  summarise(mean_energy_usage = mean(total_energy_consumption),
            in.heating_setpoint = min(in.heating_setpoint)) %>% 
  ggplot(aes(x = as.factor(in.heating_setpoint), y = mean_energy_usage)) +
  geom_boxplot(color = "purple") +
  labs(
    title = "Heating Setpoint vs. Total Energy Consumed",
    x = "Heating Setpoint (F",
    y = "Total Energy Consumed (kWh)"
  ) +
  theme_minimal()

# num occupants
energy_usage %>% 
  group_by(bldg_id) %>% 
  summarise(mean_energy_usage = mean(total_energy_consumption),
            in.occupants = min(in.occupants)) %>% 
  ggplot(aes(x = as.factor(in.occupants), y = mean_energy_usage)) +
  geom_boxplot(color = "purple") +
  labs(
    title = "Number of Occupants vs. Total Energy Consumed",
    x = "Number of Occupants",
    y = "Total Energy Consumed (kWh)"
  ) +
  theme_minimal()

# num bedrooms
energy_usage %>% 
  group_by(bldg_id) %>% 
  summarise(mean_energy_usage = mean(total_energy_consumption),
            in.bedrooms = min(in.bedrooms)) %>% 
  ggplot(aes(x = as.factor(in.bedrooms), y = mean_energy_usage)) +
  geom_boxplot(color = "purple") +
  labs(
    title = "Number of Bedrooms vs. Total Energy Consumed",
    x = "Number of Bedrooms",
    y = "Total Energy Consumed (kWh)"
  ) +
  theme_minimal()

# temperature
all_data %>% 
  mutate(month_date = format(datetime, "%b %d")) %>% 
  group_by(bldg_id, month_date) %>% 
  summarise(mean_energy_usage = mean(total_energy_consumption),
            temp = mean(`Dry Bulb Temperature [°C]`)) %>% 
  ggplot(aes(x = temp, y = mean_energy_usage)) +
  geom_point(color = "purple") +
  labs(
    title = "Temperature vs. Total Energy Consumed",
    x = "Avg Daily Temperature (C)",
    y = "Avg Daily Energy Consumed (kWh)"
  ) +
  theme_minimal()

# humidity
all_data %>% 
  mutate(month_date = format(datetime, "%b %d")) %>% 
  group_by(bldg_id, month_date) %>% 
  summarise(mean_energy_usage = mean(total_energy_consumption),
            hum = mean(`Relative Humidity [%]`)) %>% 
  ggplot(aes(x = hum, y = mean_energy_usage)) +
  geom_point(color = "purple") +
  labs(
    title = "Humidity vs. Total Energy Consumed",
    x = "Avg Monthly Humidity (%)",
    y = "Avg Monthly Energy Consumed (kWh)"
  ) +
  theme_minimal()
# by county
energy_usage %>% 
  group_by(in.county) %>% 
  summarise(tot_energy = sum(total_energy_consumption),
            n_bldgs = n_distinct(bldg_id),
            energy_per_bldg = tot_energy / n_bldgs) %>% 
  # Create the treemap
  treemap(., 
          index = c("in.county"), # Group by county
          vSize = "tot_energy",   # Size of the treemap sections
          vColor = "tot_energy",  # Color by total energy consumption
          draw = TRUE,            # Draw the treemap
          palette = "Blues",      # Set color palette
          title = "Total Energy Consumption by County")

# table
energy_usage %>% 
  group_by(in.county) %>% 
  summarise(tot_energy = sum(total_energy_consumption),
            n_bldgs = n_distinct(bldg_id), # pointless on sample of 5 houses per county
            energy_per_bldg = tot_energy / n_bldgs) %>% 
  arrange(desc(energy_per_bldg))


# by month
energy_usage %>% 
  mutate(month = month(time, label = TRUE)) %>% 
  drop_na(month) %>% 
  group_by(month) %>% 
  summarise(tot_energy = sum(total_energy_consumption)) %>% 
  # Create the treemap
  treemap(., 
          index = c("month"), # Group by county
          vSize = "tot_energy",   # Size of the treemap sections
          vColor = "tot_energy",  # Color by total energy consumption
          draw = TRUE,            # Draw the treemap
          palette = "Blues",      # Set color palette
          title = "Total County Energy Consumption by Month")

# table
energy_usage %>% 
  mutate(month = month(time, label = TRUE)) %>% 
  drop_na(month) %>% 
  group_by(month) %>% 
  summarise(tot_energy = sum(total_energy_consumption)) %>% 
  arrange(desc(tot_energy))

# by time
energy_usage %>% 
  mutate(time = format(time, "%H:%M:%S")) %>% 
  drop_na(time) %>% 
  group_by(time) %>% 
  summarise(tot_energy = sum(total_energy_consumption)) %>% 
  # Create the treemap
  treemap(., 
          index = c("time"), # Group by county
          vSize = "tot_energy",   # Size of the treemap sections
          vColor = "tot_energy",  # Color by total energy consumption
          draw = TRUE,            # Draw the treemap
          palette = "Blues",      # Set color palette
          title = "Total County Energy Consumption by Time")

# table
energy_usage %>% 
  mutate(time = format(time, "%H:%M:%S")) %>% 
  drop_na(time) %>% 
  group_by(time) %>% 
  summarise(tot_energy = sum(total_energy_consumption)) %>% 
  arrange(desc(tot_energy))

# visualize dependent variable distribution
all_data %>% 
  select(total_energy_consumption) %>% 
  ggplot(aes(x = total_energy_consumption)) +
  geom_density()
```

# Establish model predictors

```{r}
predictors <- c("month", "time", "in.sqft", "in.occupants", "in.bedrooms", 
                "Dry Bulb Temperature [°C]", "Relative Humidity [%]",
                "in.county", "in.cooling_setpoint", "in.heating_setpoint")
```

# Create model data

```{r}
all_data %>% 
  select(all_of(predictors), total_energy_consumption) %>% 
  mutate(in.occupants = as.numeric(gsub("[^0-9]", "", in.occupants)),
         in.cooling_setpoint = as.numeric(gsub("[^0-9]", "", in.cooling_setpoint)),
         in.heating_setpoint = as.numeric(gsub("[^0-9]", "", in.heating_setpoint))) %>% 
  # remove missing observations
  drop_na() -> model_data
# split into train/test datasets
train_ind <- createDataPartition(model_data$total_energy_consumption, p = 0.5) %>% .[[1]]
train <- model_data[train_ind, ]
test <- model_data[-train_ind, ]
```

# Run model

```{r}
# run models
model <- lm(total_energy_consumption ~ ., data = train)
model2 <- lm(log(total_energy_consumption) ~ ., data = train)
model3 <- lm(log(total_energy_consumption) ~ ., data = train %>% select(-in.county))
# observe model diagnostics
stargazer(model, model2, model3, 
          type = "html", 
          title = "Energy Model Comparison",
          style = "default",
          out = "model_comparison.html")
# plots
# plot(model, which = 1) # residual plot
# plot(model, which = 2) # normal q-q plot
# plot(model, which = 3) # scale-location plot (check for homoscedasticity)
# plot(model, which = 4) # Cook's distance plot (influence of each data point)
```

# Evaluate model on test data

```{r}
# establish true labels
y_true <- test$total_energy_consumption
# create predictions
predictions <- predict(model, test) %>% unname() # remove indices
predictions2 <- predict(model2, test) %>% unname() %>% exp() # remove indices
predictions3 <- predict(model3, test) %>% unname() %>% exp() # remove indices
# create validation df
data.frame(
  pred = predictions,
  pred2 = predictions2,
  pred3 = predictions3,
  true = y_true
) %>% 
  rowwise() %>% 
  mutate(
    resid = true - pred, # residuals
    resid2 = true - pred2, # residuals
    resid3 = true - pred3 # residuals
  ) -> val_df
# full model evaluation
data.table(
  "Model 1" = c(MAE(predictions, y_true),
                RMSE(predictions, y_true))
)
MAE(predictions, y_true) # MAE
RMSE(predictions, y_true) # RMSE
# residual plot on test set
val_df %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(color = "blue") +
  geom_hline(yintercept = 0, color = "red", linewidth = 3) +
  labs(
    title = "Model 1 Test Residuals",
    x = "Predicted Values",
    y = "Residuals"
  ) +
  theme_minimal()
```

Notes:

Running the linear model without log-transforming the dependent variable results in
an inability to accurately predict outliers, as the energy consumption variable is
highly right-skewed.

Run a Bayesian regression and compare validation metrics to other linear models

Ignore everything else below.

# Bayesian Regression model

```{r}
# rename columns to create valid column names for stan code
N <- nrow(train)
train_list <- list(
  N = N,
  month = as.integer(train$month),
  time = as.integer(as.factor(train$time)),
  sqft = train$in.sqft,
  occ = train$in.occupants,
  bed = train$in.bedrooms,
  temp = train$`Dry Bulb Temperature [°C]`,
  hum = train$`Relative Humidity [%]`,
  # county = as.integer(train$in.county), # remove county for the bayesian regression
  cool = train$in.cooling_setpoint,
  heat = train$in.heating_setpoint,
  total_energy_consumption = train$total_energy_consumption
)
# Stan model code as a string
stan_code <- "
data {
  int<lower=0> N;
  int month[N];
  int time[N];
  vector[N] sqft;
  vector[N] occ;
  vector[N] bed;
  vector[N] temp;
  vector[N] hum;
  vector[N] cool;
  vector[N] heat;
  vector[N] total_energy_consumption;
}

parameters {
  real alpha;                       
  real beta_month[12];                  
  real beta_time[24];                   
  real beta_sqft;                
  real beta_occ;           
  real beta_bed;            
  real beta_temp;   
  real beta_hum;      
  real beta_cool;    
  real beta_heat;    
  real<lower=0> sigma;              
}

model {
  vector[N] mu;

  // Linear predictor (sum of coefficients times the respective variables)
  for (n in 1:N) {
    mu[n] = alpha + 
            beta_month[month[n]] +      // Month-specific effect
            beta_time[time[n]] +        // Time-specific effect
            beta_sqft * sqft[n] +       // Square footage effect
            beta_occ * occ[n] +         // Occupant effect
            beta_bed * bed[n] +         // Bedroom effect
            beta_temp * temp[n] +       // Temperature effect
            beta_hum * hum[n] +         // Humidity effect
            beta_cool * cool[n] +       // Cooling setpoint effect
            beta_heat * heat[n];        // Heating setpoint effect
  }

  // Gamma likelihood for the total energy consumption
  total_energy_consumption ~ gamma(alpha, sigma * exp(mu));
}
"
# initialize values
init_fun <- function() {
  list(alpha = 1,   # Set alpha to a positive value
       beta_month = rep(0, 12),  # Initial values for beta_month
       beta_time = rep(0, 24),   # Initial values for beta_time
       beta_sqft = 0,
       beta_occ = 0,
       beta_bed = 0,
       beta_temp = 0,
       beta_hum = 0,
       beta_cool = 0,
       beta_heat = 0,
       sigma = 1)  # Set sigma to a positive value
}
# Fit the model
fit <- stan(model_code = stan_code, data = train_list, init = init_fun)

# Print the summary of the model fit
print(fit)
```

